Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increases Repair Time?

### 2015-02-26 data.tex
kinda v0.1 semi-coherent

\documentclass{sig-alternate}
\usepackage{multirow}
\usepackage{color}
\usepackage{colortbl}

\usepackage{picture}
\usepackage{wrapfig}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}} 
\newcommand{\lucas}[1]{\textcolor{red}{LUCAS: #1}} 
\newcommand{\bill}[1]{\textcolor{blue}{BILL: #1}} 
\newcommand{\carter}[1]{\textcolor{cyan}{CARTER: #1}} 
%\newenvironment{changed}{\par}{\par}

%timm tricks
\newcommand{\bi}{\begin{itemize}[leftmargin=0.4cm]}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\S\ref{sect:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}

\usepackage[shortlabels]{enumitem} 
\usepackage{times}

\usepackage{cite}
\newcommand{\subparagraph}{}
\usepackage{url}
\def\baselinestretch{1}

\setlist{nosep}

\usepackage{colortbl}
 \usepackage[font={small}]{caption, subfig}
\setlength{\abovecaptionskip}{1ex}
 \setlength{\belowcaptionskip}{1ex}

 \setlength{\floatsep}{1ex}
 \setlength{\textfloatsep}{1ex}
\usepackage[compact,small]{titlesec}
\DeclareMathSizes{7}{7}{7}{7} 
\pagenumbering{arabic}
\setlength{\columnsep}{7mm}

\begin{document}

\conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increases  Repair Time?}
\numberofauthors{3}
\author{
\alignauthor
Tim Menzies, \\Carter Pape\\
       \affaddr{CS, NcState, USA}\\
       tim.menzies@gmail.com,\\carterpape@gmail.com
\alignauthor
William Nichols,\\ Forrest Schull\\
\affaddr{SEI, CMU, USA}
wrn,fjshull@sei.cmu.edu
\alignauthor
Lucas \\Layman\\
       \affaddr{Fraunhofer Center,  USA}\\ 
       llayman@fc-md.umd.edu
}

\maketitle
\begin{abstract}
Does
repair time increases dramatically
the longer a defect persists in a system?
This  is very widely belief that is used to justify 
radical changed to    traditional software development
 processes.

This paper shows that this belief is maintained
quite strongly in the industrial software engineering
community (but, in the academic community, somewhat less so).
Yet based on a sample of 
171 software projects from around the world from 
\bill{2005 to 2013}, we argue that this belief is mostly 
incorrect. Specifically: the median fix time for issues
is just a few minutes and this does not tend to increase
the longer the issue remains in the system.

This result raises the question:
how did myths like phase delay survive so long  without  critical reassessment.  
As a community, we need to reflect more on 
research practices and on how (and when) we reflect on the prominent hypotheses of our field.
\end{abstract}

% A category with the (minimum) three required fields
\vspace{1mm}
\noindent
{\bf Categories/Subject Descriptors:} 
D.2.8 [Software Engineering]: Process metrics.

\vspace{1mm}
\noindent
{\bf Keywords:} phase delay, time to fix, agile, waterfall.

\section{Introduction}
In 2013-2014, the time of 
eleven  million programmers~\cite{pettey14} and
half a trillion dollars~\cite{avram14} was spent on information technology.
Such a large and growing effort should be managed and optimized via  well-researched conclusions.  
It is standard practice
in other fields such as medicine,
to continually revisit old conclusions~\cite{prasad13}. Software engineering should do the same lest
ill-founded  exert an undue influence on how we build software.

Accordingly, this paper revisits
the {\em phase delay} effect
i.e.  the time required to resolve an issue in a software project increases dramatically 
the longer it remains in the system. 
Phase delay has long been
a    motivator for major changes to
SE practices. Basili \& Boehm comment that since the 1980s,  phase delay:
\begin{quote}
...has been a major driver in focusing
industrial software practice on thorough
requirements analysis and design,
on early verification and validation, and
on up-front prototyping and simulation
to avoid costly downstream fixes~\cite{boehm01}.
\end{quote}
Basili and Boehm are not alone in this assessment of the significance of the phase delay effect. 
Our surveys of  industrial practitioners (shown below) reveal it to be
the single most strongly held belief amongst commercial SE engineers;
Also,
as discussed later
in this paper, concerns over phase delay were a major driver in the development of agile software
processes.

The conclusions of this paper is that it is erroneous to believe in phase delay.
When we looked for supporting evidence for that effect, we found that:
\be 
\item The evidence for that effect in the historical literature is very weak (to say the least);
\item  In  a large sample of   contemporary software projects, there is no evidence for phase delay.
 \ee
The rest of this paper present evidence for the above two points. 
The first point is  documented via  a literature survey and the second point is demonstrated
via an analysis of  
 171 software  projects developed in the period \bill{2004 to 2014}.  To the best of our knowledge,
 this is the largest study on phase delay yet seen in the literature.

The conclusions of this paper are two fold. Firstly, the SE community needs to stop
justifying changes to SE
practices via the phase delay effect. Secondly,   myths like phase delay should not have lasted decades
in the literature without  critical reassessment.  As a community, we need to reflect more on our
research practices and on how (and when) we reflect and revisit the prominent hypotheses of our field.

\section{Motivation}
\subsection{Why Reassess Existing Triusms?}

Before exploring one particular truism, we digress to discuss the merits of revisiting
old conclusions in software engineering.

Some researchers   adamantly believe in general
principles of software
engineering. Processional societies assume such generalities exist when the offer
 lists of  supposedly general ``best practices'' such as
the IEEE 1012 standard for software verification~\cite{1012}.  Endres \& Rombach's comment that 
\begin{quote}
Based on repeated and consistent observations, key lessons of these fields can now be formulated into rules or even laws, providing initial building blocks towards a theoretical foundation that is essential for further research, for teaching and for the practice of software development~\cite{endres03,rombach11}
\end{quote}
 Endres \& Rombach offer dozens of supposedly general lessons of software engineering~\cite{rombach11}.
 Many other 
widely-cited researchers  do the same; e.g.
Glass~\cite{glass02}; Jones~\cite{jones07}; Boehm~\cite{boehm00b}.
Budgen \& Kitchenham seek to reorganize SE research using
general
conclusions drawn from a larger number of studies~\cite{kitch04,budgen09}.

Nevertheless, there are many empirical  findings that 
cast doubt on the existence of  ``key lessons  that can be formulated
as general laws of SE'':
\bi
\item
Turhan~\cite{me12d} lists 28 studies with contradictory conclusions
on the relation of OO measures to defects.  Those results
 directly  contradict some of the laws listed by 
Endres \& Rombach~\cite{endres03}.

\item
Ray et al.~\cite{ray2014lang} tested if   strongly typed languages
predict for better code quality. In  728 projects,
they found  only a modest benefit in strong typing
(and caution that that effect may actually be due to other  conflating factors).
\item
Fenton \& Neil~\cite{fenton00,fenton00b}   critique the truism that
``pre-release fault rates for software
are a predictor for post-release failures'' (as claimed by~\cite{dunsmore88},
amongst others). For the systems described in~\cite{fenton97}, they
show that software modules that were highly fault-prone
prior to release revealed very few faults after release.
\item
Meyer claims that   object-oriented (OO) encapsulation will
reduce error rates in software~\cite{Meyer1988}.  Yet empirical results suggest
that debugging an OO program is many times harder and
longer than debugging a standard procedural program~\cite{hatton98}.
\item
A truism of visual programming is that ``visual
representations are inherently superior to mere textual representations''.
A review by Menzies suggests that the available
evidence for this claim is hardly conclusive~\cite{me00v}. 
\item Numerous recent {\em local learning} results compare single models
learned from all available data to multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,betta12,posnett11}.
A repeated result in those studies in that the local models generated the better effort
and defect predictions (better median results,
lower variance in the predictions).
 \item 
Other work studies  found dozens of single
models that performed worse that {\em committees of models}, all with different opinions,
that contribute to an {\em ensemble} prediction of software development effort~\cite{kocaguneli2012value,azhar13}.
\ei
To be fair to the software engineering community, SE is not the only
field where practitioners hang on to persistent beliefs, even if the evidence
for those beliefs is not strong.
The medical profession applies  many practices based on studies
that have been disproved. For example,
a  recent article
in the Mayo Clinic Proceedings~\cite{prasad13} found  
146 medical practices based on studies 
in year $i$, but which were  reversed by subsequent trials within years $i+10$.
Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}.
Aschwanden warns:
\begin{quote}
Cognitive biases - such as motivated reasoning (all of us want to believe that the things we do make a difference), base rate neglect (failing to pay attention to what happens in the absence of the intervention), and confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) - also influence how we process information~\cite{aschwanden15}.
\end{quote}

The cognitive issues that complicate medicine are also found in software engineering.
According to Passos et al.~\cite{passos11}, commercial developers
are all too willing to believe in general
truisms.
They  warn that developers
usually assume that the truisms they learn from a few past
projects are general to 
all their future projects:
\begin{quote}\label{q:pass}
...the past experiences were taken into account without 
much consideration for their context~\cite{passos11}.  
\end{quote}
The results of Jorgensen \& Gruschke support the findings if Passos et al. They report that 
  supposed software engineering    ``experts'' rarely use lessons
  from past projects to improve their future reasoning~\cite{jorgensen09}. 
 They note that
when the experts
  fail to revise their beliefs, this leads to poor
 conclusions and software projects  (see examples in~\cite{jorgensen09}).
%% This leads to  detrimental
%% situations where groups or 
%% individuals working on the same project have conflicting beliefs, which they never debate
%or refine or unify.

In summary, just like medicine, our field suffers when
 software engineers do  revise old beliefs.  Therefore, it is important
 to regularly  reexamine    old beliefs.
\subsection{Why Study Phase Delay?}
Due to its prominence in the field of software engineering,
this paper studies phase delay. As mentioned in the introduction, concerns about phase delay have
been used for decades to motivate major changes to software engineering practices. For example,
in 2000, Kent Beck used phase delay to motivate his call for ``extreme programming'' (a set of techniques
that evolved into contemporary agile and DevOps practices). Beck writes:
\begin{quote}
One of the universal assumptions of software engineering is that the cost of changing a program rises exponentially over time. I can remember sitting in a big linoleum-floored classroom as a college junior and seeing the professor draw on the board the curve found in \fig{curve1}.

The cost to fix a problem in a piece of software rises exponentially over time. A problem that might take a dollar to fix if you found it during requirements analysis might costs thousands to fix once the software is in production.

I resolved then and there that I would never let a problem get through to production. No sirree, I was going to catch problems as soon as possible. I would work out every possible problem in advance. I would review and crosscheck my code. No way was I going to cost my employer \$100,000.~\cite{beck00}
\end{quote}

\begin{figure}
 \includegraphics[width=3.3in]{beckB4.png}
 \caption{Projects exhibiting phase delay. From~\cite{beck00}.}\label{fig:curve1}
 \end{figure}
\begin{figure}
 \includegraphics[width=3.3in]{beckAFTER.png}
 \caption{Projects without   phase delay. From~\cite{beck00}.}\label{fig:curve2}
\end{figure}

Beck speculated that there might be another kind of project, shown in \fig{curve2},
that does no exhibit phase delay. In this next quote, Beck clearly states
the importance of avoiding phase delay (in fact, for Beck, it is {\em the} major
motivation for his work):
\begin{quote}
This (\fig{curve2}) is one of the premises of (agile). It is the technical premise of (agile). If the cost of change rose slowly over time, you would act completely differently from how you do under the assumption that costs rise exponentially. You would make big decisions as late in the process as possible, to defer the cost of making the decisions and to have the greatest possible chance that they would be right. You would only implement what you had to, in hopes that the needs you anticipate for tomorrow wouldn't come true. You would introduce elements to the design only as they simplified existing code or made writing the next bit of code simpler.~\cite{beck00}
\end{quote}
Beck makes one other comment that is relevant to this paper: 
\begin{quote}
The software development community has spent enormous resources in recent decades trying to reduce the cost of change-better languages, better database technology, better programming practices, better environments and tools, new notations.

What would we do if all that investment paid off? What if all that work on languages and databases and whatnot actually got somewhere? What if the cost of change didn't rise exponentially over time?~\cite{beck00}
\end{quote}
The results of this paper are that, in our sample of hundreds of projects, there is no evidence for phase delay.
This result is   consistent with Beck being correct: all that work on    languages and databases
have paid off and, hence, modern software projects do not suffer from the  phase delay effect.

XXX clumsy link here
That said, where our results
diverge from Beck is that we have observed this lack of phase delay in  projects that are not agile. Those
projects are described below.

\section{Related Work}

\input{lucasRelatedWork}

\section{lucasSurvey}

\input{lucas}

\section{billTSP}

\section{carterCharts}

\input{carter}

\section{Data}

\begin{figure}[!t]

%\renewcommand{\baselinestretch}{0.8}
\scriptsize
\begin{center}
\begin{tabular}{r|rrr|ll}
  Sample&\multicolumn{3}{c|}{Percentiles}\\ 
size & 25th & 50th & 75th & Phase injected & Phase removed\\\hline
57& 6& 16& 47&  BeforeDevelopment&Code\\
72& 2& 4& 11&  BeforeDevelopment&CodeInspect\\
165& 6& 18& 56&  BeforeDevelopment&Test\\
50& 6& 20& 37&  BeforeDevelopment&IntTest\\
42& 8& 28& 65&  BeforeDevelopment&SysTest\\\hline

66& 2& 6& 15&  Planning&Planning\\
41& 1& 5& 16&  Planning&ReqtsReview\\\hline

23& 0& 1& 3&  Reqts&Reqts\\
245& 6& 13& 23&  Reqts&ReqtsReview\\
289& 7& 16& 30&  Reqts&ReqtsInspect\\
32& 6& 21& 40&  Reqts&Design\\
49& 2& 7& 24&  Reqts&DesignInspect\\\hline

94& 2& 5& 10&  HLD&HLDReview\\
133& 3& 8& 19&  HLD&HLDInspect\\
33& 2& 5& 17&  HLD&Design\\\hline

37& 2& 4& 6&  Design&Design\\
455& 11& 24& 53&  Design&DesignInspect\\
218& 5& 14& 30&  Design&Code\\
166& 4& 11& 22&  Design&CodeInspect\\
542& 14& 31& 60&  Design&Test\\
67& 6& 15& 40&  Design&IntTest\\
92& 10& 25& 53&  Design&SysTest\\\hline

53& 2& 4& 9&  DesignInspect&DesignInspect\\
38& 1& 3& 9&  DesignInspect&Code\\\hline

126& 4& 12& 31&  Code&Code\\
459& 11& 24& 45&  Code&CodeInspect\\
461& 11& 26& 55&  Code&Test\\
348& 10& 26& 62&  Code&IntTest\\
230& 8& 18& 39&  Code&SysTest\\
71& 2& 11& 28&  Code&AcceptTest\\\hline

64& 2& 4& 10&  CodeInspect&CodeInspect\\
57& 2& 5& 13&  CodeInspect&Test\\\hline

110& 3& 10& 23&  Test&Test\\
66& 2& 4& 12&  Test&QualTest\\\hline

32& 1& 2& 4&  IntTest&QualTest\\
36& 1& 11& 30&  IntTest&IntTest\\
21& 1& 2& 5&  IntTest&SysTest\\
 \end{tabular}
\end{center}
\caption{Distribution of fix times seen in SEI TSP data.}
\label{fig:faw}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{rrl}
year& \# issues&\\\hline
2006 &  44 &\\
2007 &  34 &\\
2008&  288 &\rule{1mm}{2mm}\\
2009&  846 &\rule{3mm}{2mm}\\
2010& 1007 &\rule{3mm}{2mm}\\
2011& 3273 &\rule{10mm}{2mm}\\
2012&18102 &\rule{45mm}{2mm}\\
2013&20336 &\rule{50mm}{2mm}\\
2014& 3307 & \rule{10mm}{2mm}\\\cline{1-2}
Total:&47228
\end{tabular}
\end{center}
\caption{This paper studies 47,228 issues recorded 2006 to 2014.}\label{fig:years}
\end{figure}
\begin{figure*}[!t]
\begin{center}
\includegraphics[width=6in]{waterfall2.png}
\end{center}
\caption{Different approaches to software development:  waterfall and agile (bottom left).}
\end{figure*}

\begin{figure*}[!t]

\renewcommand{\baselinestretch}{0.7} 
\begin{center}
\begin{tabular}{ll|r|rl}
            &                  & Sample\\
Phase injected & Phase removed & size & \multicolumn{2}{l}{Scale up, w.r.t. to first phase}
\input{deltas}
\end{tabular}
\end{center}
\caption{50th percentile (median) scale ups  for  time to resolve issues (taken from \fig{raw}).}
\label{fig:scale}
\end{figure*}

In these results, we checked if a small number of bugs are most expensive. 
To check this:
\bi
\item
We repeated the analysis that generated \fig{scale},
but instead of looking at the 50th percentile, we displayed the scale up factors
\item 
We only
checked the Design and Code scale up results since, from \fig{scale}, it is clear these
have the most examples of longest phase delays.
\ei 
The  results are shown in \fig{scale90} and these
results are somewhat different for Design and Coding issues:
\bi 
\item For Design issues,  these have the same
general form as the 50th percentile results. That is, while it it certainly faster
to remove thing sin the phase where they are created, once we leave that phase
it does not seem to matter much how many phases we wait before fixing the issue.
\item
\item For Coding issues, the seems little impact of phase delay on the time
required to fix.
\ei

\begin{figure*}[!t]

\renewcommand{\baselinestretch}{0.7} 
\begin{center}
\begin{tabular}{ll|r|rl}
            &                  & Sample\\
Phase injected & Phase removed & size & \multicolumn{2}{l}{Scale up, w.r.t. to first phase}
\input{deltaBig}
\end{tabular}
\end{center}
\caption{90th percentile scale ups  for  time to resolve issues .}
\label{fig:scale90}
\end{figure*}

The data used in this work shirai14

The Team Software ProcessSM (TSPSM) provides a
framework that teams use to collect software process data in real
time, using a defined disciplined process.
In particular, three common features of the projects in our SEI TSP data set  are coaches,
peer review, and  personnel reviews:
\bi 
\item
A ``coach'' is the team member  authorized to submit project data
(before submission,
these coaches check the data for obvious errors);
\item 
``Personnel review'' is a technique taken  from the Personal Software
ProcessSM (PSPSM);
\item 
``Peer review'' is a standard technique in
traditional software engineering.
\ei
PSPSM encourages developers to continually make and review their personnel estimates
about their day-to-day tasks, then compare those estimates against the actual development effort.
In this way, developers can acquire a more realistic understanding of their work behaviour.
As to peer review,  Basili and Boehm write  commented in 2001~\cite{boehm01} 
that peer reviews can catch over half the defects introduced into a system.
Peer review can be conducted on any artifact generated anywhere in the software
lifecycle and can quickly be adapted to new kinds of artifacts.

\begin{figure}
\begin{center}
\scriptsize\begin{tabular}{|l@{~:~}l|}\hline
Bug Prediction Dataset &http://bug.inf.usi.ch \\
Eclipse Bug Data &http://goo.gl/tYKahN \\
FLOSSMetrics& http://flossmetrics.org \\
FLOSSMole &http://flossmole.org \\
IBSBSG& http://www.isbsg.org \\
ohloh& http://www.openhub.net \\
PROMISE &http://promisedata.googlecode.com \\
Qualitas Corpus &http://qualitascorpus.com \\
Software Artifact Repository &http://sir.unl.edu \\
SourceForge Research Data &http://zerlot.cse.nd.edu \\
Sourcerer Project &http://sourcerer.ics.uci.edu \\
Tukutuku &http://www.metriq.biz/tukutuku \\
Ultimate Debian Database &http://udd.debian.org\\\hline
\end{tabular}
\end{center}
\caption{Some repositories of software engineering data.}\label{fig:sedata}
\end{figure}
The data collected in the SEI TSP databases contains 
extensive process details
including a detailed phased breakdown showing what happened at what
phases of the lifecycle. This makes this data somewhat different to the standard
 publicly available
software projects data sets currently available to 
SE researchers. Most of the \fig{sedata} data sets have software product information
such as full source code, or summaries of static features.
A subset of that data contain issue or defect reports and/or
the time taken to build these systems.  We are unaware
of any of these having detailed phase-by-phase breakdowns of project data.

The phased used in this paper are shown in \fig{waterfall}. Note that, in that figure:
\bi 
\item
Several  phases have the same  sub-activities of {\em review} and {\em inspect}\footnote{\bill{plz distinguish review and inspect. does one usually happen first?}}
\item Testing is divided into several statges \footnote{\bill{need a one
line description of test vs qualtest vs inttest vs systemTest vs AcceptTest. I tries some words in the Key to \fig{sedata}. dont know if i got it right.}}
\ei 
Also shown in \fig{waterfall} (bottom left) is the standard definition of an agile process~\cite{boehmturner03}:
Given some
some backlog of tasks, wgeb teams complete their current tasks, they select the next task(s) to complete. That selection process may use a variety of criteria
to prioritize which  tasks are selected (for more details on that selection process, see~\cite{me09j,port08,boehmturner03}). Tasks are completed in ``sprints'' that can last hours,
days, but rarely not more than weeks. Each day meet for brief ``scrum'' sessions to assess (and possibly alter) their current progress on the goals of the sprint.  
Agile teams race to generate releases
(in the continuous release model, releases can be generated on a daily basis, or even faster).  
Experience gained from those releases informs the discussion in the daily scrums which, in turn,
can inform the team's decisions on how to select and implement the next set of tasks for next sprint.
In fact, that experience can result in changing some/all of the tasks in the current backlog.

@@

\section{Data}

The Software Engineering Institute (SEI) at
Carnegie Mellon University has collected data from organizations
that have adopted TSP. 
Our efforts focused on data collected by projects that launched
after 2009 and used an automatic data recording tool. 
Due to confidentiality restrictions, we cannot offer too many details
on those project. That said, there were less than a dozen United States
Department of Defence projects. Also, the majority of
the projects were weither web portals or banking systems in the US, South Africa, and Mexico. 
There were also some  medical devices (US, France, Japan, and Germany) and a few from a commercial 
computer-aided design system (developed by a world wide distributed team).

Data collected included time logs, defect
logs, added and modified size logs, and work
breakdown strcture logs.
The projects
were mostly small to medium, with a median duration of 46 days
and a maximum duration of 90 days. Median team size was 7
people, with a maximum of 40. Hence,  total development 
effort\footnote{
Effort =  duration*teamSize/workDaysPerYear.} for
these projects ranged from 1.3 years (median) to 14.3 years (max).

In these logs
the log records include work start time, work end time, delta
work time, and interruption time. Software engineers are often
interrupted by meetings, requests for technical help, reporting, and
so forth. These events are recorded, in minutes, as interruption
time. In this paper, when we report ``time to resolve an
issue'', we show the difference between the start and end times
of a work session, with any interruption time subtracted (the
difference in times, minus the interruptions).

A common property of real-world data sets is the presence
of noisy entries (superfluous  or spurious data). 
The level of noise can be quite high. As reported
in \cite{shepperd12}, around
10\% to 30\%
of the records in the NASA MDP defect data sets are
affected by noise. Nichols et al.~\cite{shirai14}  report that
the noise levesl in the SEI TSE data are smaller than those seen
in other data sets. They found in the SEI TSP data that:\bi 
\item
4\% of the data was incorrect such as  null values of illegal formats;
\item  2\% of the data has inconsistencies such as timestamps
where the stop time was before the start time;
\item 3\% of the data had data that was not credible
such as tasks listed in one day that took like than six hours.
\ei 
That said, certain semantic features of the SEI TSP data should be noted.
Firstly, in the current TSP collection tool, 
fix times are only the developer time for the developer walking through the phases of \fig{waterfall}.
We are currently tracking the fix time for post-release issues (e.g. those raised during  acceptance test and later
product life cycle). So far, in that post-release data,  we have not detected
a dramatic phase escalation effect (but at this time, we have nothing definitive comment on that matter).

\bill{somewhere you have one note on \underline{find} and fix times.  for this paper, we need just fix times. but is there
anything we need to fret about re \underline{find} times?}

The majority of the teams follow a more linear process within a single feature of physical component (usually no more than a week or two work in total). Unlike some agile teams, *most*TSP teams will not conflate development and unit test.

The starting point for this paper were the times required to fix 47,228 issues.
All data was collected
by the Software Engineering Institute in the period 2006 to 2014 (see \fig{years}).
To the best of our knowledge, this is the largest sample of ``time to fix issues''
yet analyzed.

This data was collected using the   methodology described by
Watts Humphrey's Team Software Process~\cite{tsp00}. TSP is an extension of Humphrey's early
work on Personnel Software Process~\cite{psp05}.  PSP encourages developers to continually make estimates
about their day-to-day tasks, then compare those estimates against the actual development effort
(in this way, developers can acquire a more realistic understanding of their work behaviour).

In all data, was collected from 172 separate projects. Given Beck's comments (above) about phase delay motivating agile processes, we take care to identify the differences of these 
projects to agile development projects. An alternate, more traditional approach is a ``macro linear''
approach such as the ``waterfall methodology'' that develops code after an extensive
pre-planning stage:

The defect typ that we studied were
\be 
\item Environment: design, compile, test, or other support system problems;
\item Interface: procedure calls and reference, I/O, user format;
\item Data: structure, content; 
\item Documentation: comments, messages;
\item Syntax: spelling, punctuation typos, instruction formats
\item Function: logic, pointers, loops, recursion, computation, function defects  
\item Checking: error messages, inadequate checks;
\item Build: change management, library, version control;
\item Assignment: package
declaration, duplicate names, scope, limits;
\item System: configuration, timing, memory.
\ee

\section*{Acknowledgements}
This work was partially funded by an National Science
Foundation grant NSF-CISE 1302169.

\clearpage
\vspace*{0.5mm}
\scriptsize
\bibliographystyle{plain}
\bibliography{refs}

\end{document}

One way to address the absence of general laws is  {\em local learning}.
Even if general laws
do not exist, then {\em there exists general methods for finding the best local laws}.
For example, one of us (Nichols) works extensively with software projects to help
them tune their local software process in order to better fit their local needs.XXX.

 Recent studies report that better predictors of the properties of
software projects (effort and defects) can be generated by first clustering
(a.k.a. stratifying or contextualizing or localizing) the data then learning
different models for each cluster~\cite{posnett11,betten14,betta12,yang11,yang13,minku13}.
Those {\em local learning} approach finds better predictions with lower variance than
models learned from all the unclustered data~\cite{me12d,me11m}.

The problem with the local learning is that they it can generate conclusions
that conflict with strongly-held beliefs of the humans members of the software development team.
It is an open issue how to handle those conflicts since it is unwise to
always reject the conclusions of local learning  or the software developers.
The local learners may be wrong due to incorrect assumptions by the data scientists who configured the learners~\cite{me11e,shull02}.
On the other hand,  the humans may be wrong due to the cognitive biases described above.
Experienced data scientists use a cyclic approach where
(a)~the collected data; and (b)~the learning methods; and (c) the goals of the inquiry are matured  using
insight offered by humans as they study the
feedback generated by the learners~\cite{Fayyad96,me11e}.

the learners' conclusions are discussed with the team in or

(which may be erroneous
due to incorrect assumptions

We come to this work since  recent results on {\em locality} by Menzies and others

XXXX In those 171 projects, we observed that it is
 fastest to fix issues within the same phase $i$ as when they are generated. But in a result that contradicts
 the phase delay effect, once an issue ``escapes'' phase $i$
 it so no more expensive to resolve an issue in phase $i+1$ than
 phase $i+2, i+3$, etc. Further, that increased effort may be quite modest.
 That is, contrary to established wisdom, once an  issue ``escapes'' a phase, there is no
 need to  rapidly retire that issue as soon as possible. 
Our conclusion discusses the implications for our field, and how we might better
propagated and monitor the wisdom of our field.

>>>>

The section presents the data that we use to conclude that phase delay is a not a major
contributor to the time required to resolve an issue.

The Software Engineering Institute (SEI) at
Carnegie Mellon University explores methods for software process improvement.
One product from that research is  
Watts Humphrey's  Team Software ProcessSM (TSPSM)~\cite{tsp00}. TSP is an extension of Humphrey's early
work on Personal Software
ProcessSM (PSPSM)~\cite{psp05}.

Since 2004, one  of use (Nichols) has been been mentoring 171 software development teams around the world
as they  deploy TSP within their organizations.  
The projects
were mostly small to medium, with a median duration of 46 days
and a maximum duration of 90 days. Median team size was 7
people, with a maximum of 40. Hence,  total development 
effort\footnote{
Effort =  duration*teamSize/workDaysPerYear.} for
these projects ranged from 1.3 years (median) to 14.3 years (max).
The majority of
the projects were either web portals or banking systems in the US, South Africa, and Mexico. 
There were also some  medical devices (US, France, Japan, and Germany),  a few from a commercial 
computer-aided design system (developed by a world wide distributed team), and less that  
a dozen United States
Department of Defence projects.
 An anonymized version of that data is available in the PROMISE repository\footnote{\carter{Need to add}} and,
 for confidentiality restrictions, we cannot offer 
further details
on those project.

\bill{can we offer any more deatils here?}

To aid in monitoring these 171 projects  Nichols  built a
data collection tool that gathers project data a central database at SEI. 
The issue types collected in the the SEI TSP data set divide into:
\be 
\item Environment: design, compile, test, or other support system problems;
\item Interface: procedure calls and reference, I/O, user format;
\item Data: structure, content; 
\item Documentation: comments, messages;
\item Syntax: spelling, punctuation typos, instruction formats
\item Function: logic, pointers, loops, recursion, computation, function defects  
\item Checking: error messages, inadequate checks;
\item Build: change management, library, version control;
\item Assignment: package
declaration, duplicate names, scope, limits;
\item System: configuration, timing, memory.
\ee
The tool also collected process information about those defects.
That data includes  work start time, work end time, delta
work time, and interruption time. Software engineers are often
interrupted by meetings, requests for technical help, reporting, and
so forth. These events are recorded, in minutes, as interruption
time. In this paper, when we report ``time to resolve an
issue'', we show the difference between the start and end times
of a work session, with any interruption time subtracted (the
difference in times, minus the interruptions).

Three common features of all those  are (1)~coaches,
(2)~peer review, and  (3)~personnel reviews.

A ``coach'' is a team member  authorized to submit project data
Before submission,
these coaches check the data for obvious errors.

``Personnel review'' is a technique taken  from the Personal Software
ProcessSM (PSPSM). PSPSM encourages developers to continually make and review their personnel estimates
about their day-to-day tasks, then compare those estimates against the actual development effort.
In this way, developers can acquire a more realistic understanding of their work behaviour.

``Peer review'' is a  technique in
traditional software engineering.
 Basili and Boehm write  commented in 2001~\cite{boehm01} 
that peer reviews can catch over half the defects introduced into a system.
Peer review can be conducted on any artifact generated anywhere in the software
lifecycle and can quickly be adapted to new kinds of artifacts.

%That said, certain semantic features of the SEI TSP data should be noted.
%Firstly, in the current TSP collection tool, 
%fix times are only the developer time for the developer walking through the phases of \fig{waterfall}.
%We are currently tracking the fix time for post-release issues (e.g. those raised during  acceptance test and %later
%product life cycle). So far, in that post-release data,  we have not detected
%a dramatic phase escalation effect (but at this time, we have nothing definitive comment on that matter).%
%
%\bill{somewhere you have one note on \underline{find} and fix times.  for this paper, we need just fix times. %but is there
%anything we need to fret about re \underline{find} times?}

\subsection{Project Phases}
This paper studies the impact of phase delay on the time required to resolve issues.
The phased used in this paper are shown in \fig{waterfall}. Note that, in that figure:
\bi 
\item
Several  phases have the same  sub-activities of {\em review} and {\em inspect}\footnote{\bill{plz distinguish review and inspect. does one usually happen first?}}
\item Testing is divided into several statges \footnote{\bill{need a one
line description of test vs qualtest vs inttest vs systemTest vs AcceptTest. I tries some words in the Key to \fig{sedata}. dont know if i got it right.}}
\ei 
Also shown in \fig{waterfall} (bottom left) is the standard definition of an agile process~\cite{boehmturner03}:
Given some
some backlog of tasks, wgeb teams complete their current tasks, they select the next task(s) to complete. That selection process may use a variety of criteria
to prioritize which  tasks are selected (for more details on that selection process, see~\cite{me09j,port08,boehmturner03}). Tasks are completed in ``sprints'' that can last hours,
days, but rarely not more than weeks. Each day meet for brief ``scrum'' sessions to assess (and possibly alter) their current progress on the goals of the sprint.  
Agile teams race to generate releases
(in the continuous release model, releases can be generated on a daily basis, or even faster).  
Experience gained from those releases informs the discussion in the daily scrums which, in turn,
can inform the team's decisions on how to select and implement the next set of tasks for next sprint.
In fact, that experience can result in changing some/all of the tasks in the current backlog.

@@
